---
title: "Machine learning with tidymodels"
subtitle: "Regression analysis of electric vehicle ranges" 
author: "Made by:Hampus Nordholm"
date: "2024-10-06"
output:
  pdf_document:
    latex_engine: xelatex
---
# Introduction 
The report aims to analyze the average effect on the range in miles based on electric vehicle type (battery electric vehicle or plug-in hybrid). A linear regression model was trained using tidymodels and evaluated on a testing split. Correlation analysis was also performed to identify features with a linear relationship to longer range. Through regression analysis, the main goal is to answer the question: 

What is the effect on range in miles for electric vehicles depending on whether they are battery electric vehicles or plug-in hybrids?

# Solution summary
The regression model achieved an RMSE of 53.8, an R-squared value of 0.705, and an MAE of 36.1 when evaluated on the testing split. There is a statistically significant difference in range in miles between plug in hybrids and battery electric vehicle types, with an average decrease of 167 miles for plug in hybrids compared to battery electric vehicles. Through correlation analysis, it was concluded that Tesla models such as the Bolt EV and Model 3 are the most strongly correlated with the range bin of 215 to infinity. Chevrolet and Tesla were the only manufacturers positively correlated with the highest range bin within this analysis.

# Core syntax for analysis 
```{r,warning=FALSE,message=FALSE}
#CORE LIBRARIES 

#Data analysis 
library(tidyverse)
library(correlationfunnel)
library(skimr)
library(janitor)

#Machine learning
library(tidymodels)
```

```{r}
#Loading data -- 
electric_tbl <- read_csv("data.csv")
```


```{r}
#Data exploration --

electric_tbl %>% glimpse()

electric_tbl %>% sample_n(20)

#Cleaning var names -- 
electric_tbl <- electric_tbl %>% clean_names()

```

```{r}
#EXPLORATORY DATA ANALYSIS --

# Count n vehicles where distance = 0  --
electric_tbl %>% 
  filter(electric_range==0) %>% 
  count()

# Basic mean values for  electric range by EV type -- 

electric_tbl %>% 
  filter(electric_range!=0) %>% 
  group_by(e_v_type) %>% 
  summarise(avg_electric_range=mean(electric_range)) %>% 
  ggplot(aes(e_v_type,avg_electric_range,fill=e_v_type))+
  geom_col()+
  geom_text(aes(label=round(avg_electric_range,1)),
            vjust=-0.5)+
  labs(title="Avg range for respective EV type",
       x="Type",y="Miles")

#Histogram for respective group -- 
electric_tbl %>% 
  filter(electric_range!=0) %>% 
  ggplot(aes(electric_range,fill=e_v_type))+
  geom_histogram(alpha=0.7,bins=25)+
  labs(title="Range distribution for EV-type",
       x="Range(miles)")

```

```{r}
#Correlation analysis -- 

electric_tbl %>% 
  filter(electric_range!=0) %>% 
  na.omit() %>% 
  binarize() %>% 
  correlate(electric_range__215_Inf) %>% 
  plot_correlation_funnel()

```


```{r}
# Simple linear reg feature selection -- avg. electric range effect of EV-type**

ev_type_tbl <- electric_tbl %>%
  filter(electric_range!=0) %>% 
  select(electric_range,e_v_type) %>% 
  mutate(e_v_type=as.factor(e_v_type))

# Train / test split 
set.seed(123)
simple_lm_split <- initial_split(data=ev_type_tbl,prop=0.8)
lm_training <- training(simple_lm_split)
lm_testing <- testing(simple_lm_split)

#Regression recipe -- 
lm_model_rec  <- recipe(electric_range~e_v_type,data=lm_training)

# Linear model spec -- 
lm_model_spec<-linear_reg() %>% 
  set_engine("lm")

#Combind into workflow -- 

lm_wf <- workflow() %>% 
  add_recipe(lm_model_rec) %>% 
  add_model(lm_model_spec)

# Training linear model -- 

lm_model_fit <- fit(lm_wf,data=lm_training)

#Results -- 
lm_model_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

#Model evaluation -- (Examined on testing data)
ev_predict <- predict(lm_model_fit,new_data=lm_testing)

#Combinding actual vs. predicted values -- 
actvspred_lm<- lm_testing %>% select(electric_range) %>% 
  bind_cols(ev_predict)

lm_evaluation <-metrics(data=actvspred_lm,truth=electric_range,estimate=.pred)

#Final linear reg. model metrics -- 
lm_evaluation

```








